---
title: "Programming of Data Science"
author: "Garima Mahajan"
date: "21/01/2019"
output: word_document
---
4.1 Validating the Strike Rate

# Creating the data frame
# the data frame command is used for the formulation of the required dataframe.
#The dataframe takes seven variables namely: Batsman, `Hows Out`, Runs, Balls, `4s`, `6s`, `SR`. 


```{r}
cric.df <- data.frame(
  Batsman = c("KL Rahul", "M Vijay", "CA Pujara", "V Kohli (c)", 
              "AM Rahane", "RG Sharma", "RR Pant +", "R Ashwin", 
              "I Sharma", "Mohammed Shami", "JJ Bumrah"),
  `How Out` = c("c Finch b Hazlewood", "c +Paine b Starc", "run out (Cummins)",
                "c Khawaja b Cummins", "c Handscomb b Hazlewood", "c Harris b Lyon",
                "c +Paine b Lyon", "c Handscomb b Cummins", "b Starc",
                "C +Paine b Hazlewood", "not out"),
  Runs = c(2, 11, 123, 3, 13, 37, 25, 25, 4, 6, 0),
  Balls = c(8, 22, 246, 16, 31, 61, 38, 76, 20, 10, 0),
  `4s` = c(0, 1, 7, 0, 0, 2, 2, 1, 1, 1, 0),
  `6s` = c(0, 0, 2, 0, 1, 3, 1, 0, 0, 0, 0),
  `SR` = c(25.00, 50.00, 50.00, 18.75, 41.93, 60.65, 65.78, 32.89, 20.00, 60.00, NA)
)
```

```{r}
print(cric.df)
```
# the formula for calculating the strike rate is the ratio of runs and balls multiplied by hunderd.

```{r}
cric.df$SR_New <- (cric.df$Runs/cric.df$Balls)*100
```
#calculating the error

```{r}
cric.df$error <- cric.df$SR_New - cric.df$SR
```

#rounding the error
```{r}
cric.df$error <- round(cric.df$error, digits = 2)
print(cric.df)
```

# It is seen that the error is observed in row 5, 6 and 7. 

4.2 Ball Outcome 

# Declaring the state variable as list
```{r}
state = list (balls = 0, runs = 0, fours = 0, sixes = 0, extras = 0, out = FALSE)
```

# formulating oneBall function with multiple if-else statements. 
#The variable state and outcome has been used as arguments to function: oneBall

```{r}
oneBall <- function(state, outcome){
  if (outcome == "0 runs"){
    state$balls = state$balls + 1
    } else if (outcome == "1 run"){
    state$balls = state$balls + 1
    state$runs = state$runs + 1
  } else if (outcome == "2 runs"){
    state$balls = state$balls + 1
    state$runs = state$runs + 2
  } else if (outcome == "4 runs"){
    state$balls = state$balls + 1
    state$runs = state$runs + 4
    state$fours = state$fours + 1
  } else if (outcome == "6 runs"){
    state$balls = state$balls + 1
    state$runs = state$runs + 6
    state$sixes = state$sixes + 1
  } else if (outcome == "Wide"){
    state$extras = state$extras + 1
  } else if (outcome == "No ball"){
    state$extras = state$extras + 1
  } else if (outcome == "Out"){ 
    state$out=TRUE
    state$balls = state$balls + 1
    }
  print(state)
}

oneBall(state, "6 runs")

```

4.3 Batsman’s score
# Formulating oneBatsman function by using similar control statements but by taking only one argument i.e. state.

```{r}
state = list(balls = 0, runs = 0, fours = 0, sixes = 0, extras = 0, out = FALSE)

```

```{r}
oneBatsman <- function(state){
  repeat{
    outcome = sample(c("0 runs", "1 runs", "2 runs", "4 runs", "6 runs", "Wide", "No Ball", "Out"), size = 1)
    
state = oneBall(state, outcome)
    if (outcome == "Out") {
      return(state)	#the function returns the updated state
      print(state)	#the function also prints the updated states
      break 		    #the repeat loop is broken when outcome is out
    }
  }
}

```

```{r}
oneBatsman(state)  
```

4.4 Team Score

#creating a simulation for 10 batsman and storing the data into the dataframe d


```{r}
d = NULL
for (player in 1:10) {
  state = list(balls = 0, runs = 0, fours = 0, sixes = 0, extras = 0, out = FALSE)
  u_state = oneBatsman(state)
  u_state = data.frame(u_state)
  u_state$player = player
  d = rbind(d, u_state) #append the data to new database
}
print(d)
```

```{r}

```


#getting total runs by adding runs and extras
```{r}
`total runs` = sum(d$runs)+sum(d$extras)

print(`total runs`)

```


